Pesticides pollution of small streams in Germany

### 2016-07-25 article.tex
some changes

More than 50\% of the total land area in Germany are used by agriculture \citep{statistisches_bundesamt_bodenflache_2014}.
In the year 2014 more the 45000 tonnes of 766 authorized pesticides were sold for application on these areas \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
The applied pesticides may enter surface waters via spray-drift, edge-off-field run-off or drainage, with run-off being one of the major input routes \citep{schulz_comparison_2001}.
Once entered the surface waters pesticides are frequently detected in environmental monitoring \citep{malaj_organic_2014} and may have adverse effects on biota and ecosystem functioning \citep{schulz_field_2004, schafer_effects_2007}.

% Small waterbodies

% runoff during rainfall (Liess et al. 1999)

% underestiamtion of grab sampling (Stehle et al. 2013).

The aim of this study was 
(i) to compile monitoring data on a national scale and to answer the questions: \\
(ii) Is the data a representative description of the pollution situation? \\
(iii) Are small agricultural waters more polluted compared to bigger streams? Are there thresholds? \\
(iv)  How polluted are small streams and which pesticides are responsible?

@@

We homogenized and unified these datasets into a common database.
We implemented a robust and transparent data cleaning work flow \citep{poisot_best_2015}, though parts of the dataset are proprietary.
An overview of the data cleaning process is provided in the supplemental materials.  
The compiled dataset comprised only a small fraction of standing waters and most of the samples where sampled via grab sampling.
Therefore, we report  only results for grab sampling from streams.  \todo{move to results and say that nearly no information on standing waters is available}
To assess whether  grab samples were taken during potential rainfall events we intersected sampling coordinates with daily precipitation data \citep{rauthe_central_2013} from the sampling date and the day before .

@@

\begin{enumerate}
\item National and international Environmental Quality Standards (EQS) \citep{ogewv_verordnung_2011,european_union_directive_2013}:
We used only Maximum Annual Concentration EQS (MAC-EQS) for characterization.

\item Regulatory Acceptable Concentrations (RAC) \citep{brock_linking_2010}:
This is the lowest concentration at which no acceptable biological effects are expected. 
These are derived during authorization  process of pesticides and contain an uncertainty factor.
The German Federal Environmental Agency provided RACs for this study.  
We expressed  RAC as Risk Quotient (RQ)

@@

We compiled a national scale dataset  comprising 42236 samples from 3049 sampling sites (Figure \ref{fig:fig1} and Supplement). 
We found big differences in the number of sampling sites between federal states.

@@

Vergleich mit der Schweiz.....

>>>>

More than 50\% of the total land area in Germany are used by agriculture \citep{statistisches_bundesamt_bodenflache_2014}.
In the year 2014 more the 45,000 tonnes of 766 authorized pesticides were sold for application on this area \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
The applied pesticides may enter surface waters via spray-drift, edge-off-field run-off or drainage, with run-off being one of the major input routes \citep{schulz_comparison_2001,liess_determination_1999}.
Once entered the surface waters pesticides are frequently detected in environmental monitoring \citep{malaj_organic_2014} and may have adverse effects on biota and ecosystem functioning \citep{schulz_field_2004, schafer_effects_2007}.

National monitoring programs are setup for determination and surveillance of the chemical and ecological status of surface, ground and drinking water.
These monitoring programs produce huge amounts of data, which possibly can also be used to answer other questions.
In Germany monitoring programs are setup by the federal states in compliance with the Water Framework Directive \citep{quevauviller_water_2008} and further state specific needs.
However, currently there is no curated national-wide compilation of this data.

Small water bodies are important refuges of biodiversity \citep{davies_comparison_2008} and enabling downstream colonisation of 
At the same time they may be exposed to a high risk of pesticide contamination from adjacent agricultural areas and low dilution effects \citep{liess_determination_1999}.

% Although small waterbodies make up most

% running  water  
% with  at  least  10km2  catchment  area

% However, very little is know about the pollution of such streams. 
% The

% Small waterbodies
% underestimation of grab sampling (Stehle et al. 2013).

The aim of this study was to compile monitoring data on a national scale and to answer the questions: \\
(i) Is the currently available monitoring data a representative description of the pollution situation? \\
(ii) Are small agricultural waters more polluted compared to bigger streams? Are there thresholds? \\
(iii)  How polluted are small streams and which pesticides are responsible?

@@

We homogenized and unified these datasets into a common database.
We implemented a robust and transparent data cleaning work flow \citep{poisot_best_2015}, though parts of the dataset are proprietary.
An overview of the data cleaning process is provided in the supplemental materials.  
To assess whether samples were taken during potential rainfall events we intersected sampling coordinates with daily precipitation data \citep{rauthe_central_2013} from the sampling date and the day before.

@@

\begin{enumerate}
  \item National and international Environmental Quality Standards (EQS) \citep{ogewv_verordnung_2011,european_union_directive_2013}:
  We used only Maximum Annual Concentration EQS (MAC-EQS) for characterization.
  \item Regulatory Acceptable Concentrations (RAC) \citep{brock_linking_2010}:
  This is the lowest concentration at which no acceptable biological effects are expected. 
  These are derived during authorization  process of pesticides and contain an uncertainty factor.
  The German Federal Environmental Agency provided RACs for this study.  
  We expressed  RAC as Risk Quotient (RQ)

@@

The compiled dataset comprised only a small fraction of standing waters and most of the samples where sampled via grab sampling. \todo{zahlen}
Therefore, we report only results for grab sampling from streams.

The analysed dataset comprised 42236 samples from 3049 sampling sites (Figure \ref{fig:fig1} and Supplement). 
We found big differences in the number of sampling sites between federal states.

@@

% Vergleich mit der Schweiz.....

%%% Risk assessment
% findings from field studies on pesti-
% cide exposure and effects must be used for a retrospec-
% tive validation of the current EU regulatory risk assess-
% ment, particularly for its future development.

% validation of the risk assessment through
% targeted chemical and biological post-authorization
% monitoring programs must be implemented for com-
% pounds of concern to ensure that their application
